### 2019-01-23 journal/abstract.tex
Example is ready to go

Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.  This paper proposes a metric for measuring distance between executions, and specializes it to an important application: determining similarity of failing test cases for the purpose of automated fault identification and localization in debugging based on automatically generated compiler tests.  The metric is based on a causal concept of distance where executions are similar to the degree that changes in the program itself, introduced by mutation, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases (for the original compiler) become successful due to the same mutant, they are more likely to be due to the same fault.  We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.

>>>>

Measuring the distance between two program executions is a fundamental problem in dynamic analysis of software, and useful in many test generation and debugging algorithms.  This paper proposes a metric for measuring distance between executions, and specializes it to an important application: determining similarity of failing test cases for the purpose of automated fault identification and localization in debugging based on automatically generated compiler tests.  The metric is based on a causal concept of distance where executions are similar to the degree that changes in the program itself, introduced by mutation, cause similar changes in the correctness of the executions.  Specifically, if two failing test cases for a compiler can be made to pass by applying the same mutant, those two tests are more likely to be due to the same fault.  We evaluate our metric using more than 50 faults and 2,800 test cases for two widely-used real-world compilers, and demonstrate improvements over state-of-the-art methods for fault identification and localization.  We additionally show that this approach, though devised for compilers, is applicable as a conservative fault localization algorithm for other types of programs.

### 2019-01-23 journal/example.tex
Example is ready to go

\begin{figure}
\begin{code}
Test 1:  (1, 2)
Test 2:  (3)
Test 3:  (5, 6)
\end{code}
\caption{Failing tests for the toy compiler.}
\label{fig:toybugs}
\end{figure}

@@

Assume that all of our test cases run the compiler with {\tt optimizeLevel} of 2, so as to potentially apply all optimizations.  A popular way to fuzz a compiler is to compare program behavior when code is compiled with all optimizations with code compiled with no optimizations, so this reflects real-world compiler testing well.  A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.

Further, let us only consider two kinds of mutants:  condition-negation and statement deletion mutants.  Obviously, negating the condition on line 3 will make all of our failing test cases pass, and negating the condition on line 7 will make test cases 2 and 3 pass, but not test case 1.  Negating the conditions on lines 4-5, 8-9, and 11-12 will repair our test cases, but is also likely to make some passing test case fail, so these mutants will not be used in our analysis.  Finally, the only statements that can be deleted without breaking the code are deletions of the {\tt corrupt} calls on lines 6, 10, and 13.  We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.  Each test can then be associated with a vector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.

>>>>

\begin{figure}
\begin{code}
T1: (1, 2)
T2: (3)
T3: (5, 6)
\end{code}
\caption{Failing tests for the toy compiler.}
\label{fig:toybugs}
\end{figure}

@@

Assume that all of our test cases run the compiler with {\tt
  optimizeLevel} of 2, so as to potentially apply all optimizations.
A popular way to fuzz a compiler is to compare program behavior when
code is compiled with all optimizations with program behavior when
code is compiled with no optimizations, so this reflects real-world compiler testing.  A set of delta-debugged failing tests, then, might be those shown in Figure \ref{fig:toybugs}.

Further, let us only consider two kinds of mutants:
condition-negation and statement-deletion mutants.  Obviously,
negating the condition on line 3 will make all of our failing test
cases pass, and negating the condition on line 7 will make test cases
2 and 3 pass, but not test case 1.  Negating the conditions on lines
4-5, 8-9, and 11-12 will repair our test cases, but these changes are
also likely to make some passing test case fail (they represent
preconditions for compiler optimizations), so these mutants will not
be used in our analysis.  Finally, the only statements that can be deleted without breaking the code are deletions of the {\tt corrupt} calls on lines 6, 10, and 13.  We can identify these mutants, which will form the basis of our distance metric and our fault localization, by the line number of the change, since there is only one mutant per line in the set.  Each test can then be associated with a bitvector of 5 elements:  {\tt (repaired by line 3, repaired by line 6, repaired by line 7, repaired by line 10, repaired by line 13)}.  The vectors for our tests, then, are:

\begin{code}
T1: 11000
T2: 10110
T3: 10101
\end{code}

T1 thus has a distance of 0.75 from T2 and T3.  T2 has
a distance of only 0.5 from T3, however.

\subsection{Fuzzer Taming for the Toy Compiler}

If the Furthest-Point-First
algorithm starts from T1 it will arbitrarily select either T2
or T3 next, then the other test.  If it starts with T2 or T3, it
will select T1 next, then the other of the two closer tests.  Of
course, taming a set of three failing tests is not difficult!
However, if we imagine a much larger set of failing tests, which
contain irrelevant features that cannot be removed by delta-debugging
without causing the programs to become syntactically invalid, so long
as the basic structure of the cause for failure remains the same, the
the FPF algorithm will still select tests with the same bitvectors as
T1, T2, and T3.  Concretely, if we add these tests (shown with associated bitvectors):

\begin{code}
T4: (1, 2, 7)    11000
T5: (3, 6)       10110
T6: (1, 5, 6)    10101
T7: (1, 2, 5, 6) 10000
\end{code}

\noindent then FPF from T1 will proceed to a maximally distant
test, such as T5, at distance 0.75.  The next ranked test could not have
been T4 (distance 0.0) or T7 (distance 0.5), but could have
been T2, T3, or T6.  After T1 and T5, the next test in the ranking will be the test whose nearest
neighbor among already chosen tests is furthest away.  The distances
to the two tests already chosen are:

{\scriptsize
\begin{code}
                         11000  10110
T2: (3)          10110   0.75   0.00
T3: (5, 6)       10101   0.75   0.50
T4: (1, 2, 7)    11000   0.00   0.75
T6: (1, 5, 6)    10101   0.75   0.50
T7: (1, 2, 5, 6) 10000   0.50   0.67
\end{code}
}

\noindent The FPF algorithm could choose either T3, T6, or T7 all of
which have a nearest neighbor at distance 0.5.  Let
us assume it chooses T6.  At this point, the three distinct bugs
are all ranked.  However, there is one interesting pattern remaining.
The distances to tests are now:

{\scriptsize
\begin{code}
                         11000  10110  10101
T2: (3)          10110   0.75   0.00   0.50
T3: (5, 6)       10101   0.75   0.50   0.00
T4: (1, 2, 7)    11000   0.00   0.75   0.75
T7: (1, 2, 5, 6) 10000   0.50   0.67   0.67
\end{code}
}

\noindent The algorithm is forced to choose T7, the only remaining
test that does not have the same bitvector as another test.  T7 is
worth ranking, as it shows that a program can be written that will
result in wrong code unless both bugs it triggers are fixed.

\subsection{Fault Localization for the Toy Compiler}

Now let us consider the {\bf Repair} fault localization algorithm
applied to this example.  If we are trying to fix the problem
exhibited by T2, what will the system suggest to us as the
location of the fault?  Recall that 1) in {\bf Repair}, a localization
is the location (and content) of a repairing mutant.  T2 is repaired
by negating the condition on line 3, negating the condition on line 7,
or deleting the statement on line 10.  Which of these is best?  {\bf
  Repair} distinguishes between repairing mutants by ordering them by
their distance to the most-distant (by bitvector) test also repaired
by that mutant.  The only other tests repaired by deletion of line 10
are the other instances of the same bug, all at distance 0.  Of the
other two repairs, negating the condition on line 3 also repairs T1,
at distance 0.75, while negating the condition on line 2 repairs
nothing farther away than T3 at 0.5.
Therefore, the localization for T2 is (in rank order):

{\scriptsize {\tt 1: DELETE  10                corrupt(internalRepresentation);}}

{\scriptsize {\tt 2: NEGATE  07       if optimizeLevel > 1: \{}}

{\scriptsize {\tt 3: NEGATE  03    if optimizeLevel > 0: \{}}

\subsection{The Onion Ring}

In this simple toy example, the onion ring will only have a few
layers, one for each possible distance from the test at the center.
For example, if we take T2 as the center, then the layers will be:

\begin{code}
 0.00:  T2 T5
 0.50:  T3 T6
 0.67:  T7
 0.75:  T1 T4
\end{code}

### 2019-01-23 journal/introduction.tex
Example is ready to go

The proposed techniques in this paper are motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in modern optimizing compilers.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
This paper proposes the use of mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. \comment{The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.  Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).}

@@

This leads to the general idea of comparing any two executions by their response to mutants, since a mutant can cause a passing or failing test to violate more/different properties.   However, for practical purposes the most interesting implication is that if two failing test cases can be made to differ as to whether they fail by a mutant, this provides evidence that they fail due to different faults.

\comment{
%Where there are multiple, independent, correctness properties, this even allows us to compare failing and successful executions, since a mutant can cause a failure to violate a different property.
{\bf Hypothesis 2} at first may suggest that this idea subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.  The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.  This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.  Hypothesis 2 also suggests an additional hypothesis related to hypothesis 1.  This is {\bf Hypothesis 1a:} If two failing test cases can be made to differ as to whether they fail by a modification of $P$, this provides evidence that they fail due to different faults.
}

>>>>

The proposed techniques in this paper are motivated in large part by progress in automated testing.  Recent advances, some implemented in powerful tools \cite{csmith,jsfunfuzz,ISSTA12,LangFuzz,ZhendongPLDI,ZhendongOOPSLA}, have made it easier to find subtle faults in modern optimizing compilers.  Unfortunately, the blessing of many failing test cases can become a curse for a developer.  Simply discovering which test cases in a large set of failures represent distinct faults (or already known faults) is a very difficult problem \cite{PLDI13,Podgurski04}.  For simple crashes and assertion violations, bucketing failures is sometimes easy.  When a failure is detected by differential testing \cite{Differential}, where two compilers or optimization levels disagree, the problem is so hard that it may turn potential users away from automated test generation \cite{PLDI13}.  Even once a fault is identified, fixing the problem can be very difficult.  Due to complexity of intermediate representation, interaction of optimizations, subtle language semantics, and other issues, debugging compiler faults may often be harder than debugging faults in other kinds of software (this is probably not limited to compilers; debugging systems software in general is hard \cite{mickens}).
This paper proposes the use of mutants \cite{mutant} to assist debugging based on large sets of automatically generated test cases.  The underlying idea is that if two failures can be repaired in the same way, even if that repair only avoids the failure and does not truly fix it, they are likely due to the same fault. Repairs can also provide  information for understanding and debugging discovered faults. The approach suggests a plausible metric for comparing any two program executions, not just failing test cases.  Comparing executions has been a prominent concept in dynamic analysis throughout much of the last two decades \cite{NearNeighbor,Sumner2011} (e.g., Ball's early discussion of profile similarity \cite{BallConcept}).

@@

This leads to the general idea of comparing any two executions by their response to mutants, since a mutant can cause a passing or failing test to violate more/different properties.   
This generalization to any two executions may at first glance suggest that our approach subsumes past notions of similarity, such as spectra of code coverage, since if two executions do not cover a statement, they both cannot alter correctness if that statement is altered.  The metrics are in fact simply incomparable:  for example, differences in coverage that never have an effect on correctness cannot contribute to similarity in our approach; with weak correctness properties most executions are very similar.  This is similar to the idea of \emph{checked coverage}:  code that cannot impact correctness is not interesting \cite{CheckCov}.  
However, for practical purposes the most interesting implication of our proposal is that if two failing test cases can be made to differ as to whether they fail by applying a program mutant, this provides evidence that they fail due to different faults.
